### 2014-11-08 store-opsem.tex
Added proof for Soundness Modulo Classification

Figure ~\ref{sem:oper} presents operational semantics as rules
defining the reduction relation ($\xrightarrow{}$) over the execution
state.  Since we now have a concrete store, we extend our system model
with $\Theta$, a representation of the store as a map from replicas to
their local states. The local state of a replica $r$ (i.e.,
$\Theta(r)$) is a set of effects that are currently visible at $r$.
An operation $op$ performed at replica $r$ can only witness the set of
effects ($\Theta(r)$) visible at $r$. For the sake of clarity, we only
consider a single replicated object of well-defined type (for eg: a
replicated object of type \cf{BankAccount}) in our formalization.  Our
semantics are parametric over the specification of this replicated
data type. Figure~\ref{sem:oper} formalizes replicated data type (RDT)
specification as tuple $(\delta,\Ops,\Ctrts)$, where $\delta$ is the
data type, $\Ops$ maps labels ($op$) of operations on $\delta$ to
their definitions, while $\Ctrts$ maps them to their consistency
contracts ($\cv$). The definition of an operation is expected to be a
lambda expression, although we do not enforce this in our
formalization. For technical reasons, we tag each session with a
session identifier ($s$) and the sequence number ($i$) of the next
operation in the session.

@@

\begin{theorem}[Soundness Modulo Classification]
\label{thm:weak-soundness}
For every well-formed execution state $\E$, and contract class $\tau 
\in \{{\sf ec},{\sf cc},{\sf sc}\}$, If:
\begin{smathpar}
(\E, \Theta, \langle s,i,\langle op,\tau \rangle;
\sigma \rangle \pll \Sigma) \xrightarrow{\eff} (\E', \Theta',
\langle s,i,\sigma \rangle \pll \Sigma')
\end{smathpar}
then (\romannumeral 1) $\E'$ is well-formed, and (\romannumeral 2)
$\E' \models \cv_{\tau}[\eff/\cureff]$
\end{theorem}
\begin{proof}
  By case analysis on $(\E, \Theta, \langle s,i,\langle op,\tau \rangle;
  \sigma \rangle \pll \Sigma) \xrightarrow{\eff} (\E', \Theta',
  \langle s,i,\sigma \rangle \pll \Sigma')$. 
  \GK{Lets first finalize all notations, rules, and also agree on
  all theorem statements. This way we don't have to change the proof
  again.} 
\end{proof}

\begin{corollary}[Soundness]
\label{thm:soundness}
For every well-formed execution state $\E$, and contract class $\tau
\in \{{\sf ec},{\sf cc},{\sf sc}\}$, and for every consistency
contract $\cv$ in the contract class $\tau$, If:
\begin{smathpar}
(\E, \Theta, \langle s,i,\langle op,\tau \rangle;
\sigma \rangle \pll \Sigma) \xrightarrow{\eff} (\E', \Theta',
\langle s,i,\sigma \rangle \pll \Sigma')
\end{smathpar}
then (\romannumeral 1) $\E'$ is well-formed, and (\romannumeral 2)
$\E' \models \cv[\eff/\cureff]$
\end{corollary}
\begin{proof}
Follows from Theorems~\ref{thm:classification-sound} and
\ref{thm:weak-soundness}
\qed
\end{proof}

>>>>

Figure ~\ref{sem:oper} presents operational semantics as rules
defining the reduction relation ($\xrightarrow{}$) over the execution
state.  Since we now have a concrete store, we extend our system model
with $\Theta$, a representation of the store as a map from replicas to
their local states. The local state of a replica $r$ (i.e.,
$\Theta(r)$) is a set of effects that are currently visible at $r$.
An operation $op$ performed at replica $r$ can only witness the set of
effects ($\Theta(r)$) visible at $r$. To avoid too many paranthesis in
our formal presentation, we represent operation $op$, whose contract
is in consistency class $\tau$, as $op_\tau$ instead of the usual
$\langle op,\tau \rangle$. For the sake of clarity, we only
consider a single replicated object of well-defined type (for eg: a
replicated object of type \cf{BankAccount}) in our formalization.  Our
semantics are parametric over the specification of this replicated
data type. Figure~\ref{sem:oper} formalizes replicated data type (RDT)
specification as tuple $(\delta,\Ops,\Ctrts)$, where $\delta$ is the
data type, $\Ops$ maps labels ($op$) of operations on $\delta$ to
their definitions, while $\Ctrts$ maps them to their consistency
contracts ($\cv$). The definition of an operation is expected to be a
lambda expression, although we do not enforce this in our
formalization. For technical reasons, we tag each session with a
session identifier ($s$) and the sequence number ($i$) of the next
operation in the session.

@@

% First, we prove few useful lemmas:

% \begin{lemma}[]
% \end{lemma}

\begin{theorem}[Soundness Modulo Classification]
\label{thm:weak-soundness}
For every well-formed execution state $\E$, and contract class $\tau 
\in \{{\sf ec},{\sf cc},{\sf sc}\}$, If:
\begin{smathpar}
(\E, \Theta, \langle s,i,op_\tau ::
\sigma \rangle \pll \Sigma) \xrightarrow{\eff} (\E', \Theta',
\langle s,i+1,\sigma \rangle \pll \Sigma)
\end{smathpar}
then (\romannumeral 1) $\E'$ is well-formed, and (\romannumeral 2)
$\E' \models \cv_{\tau}[\eff/\cureff]$
\end{theorem}
\begin{proof}
  By case analysis on the derivation:
  \begin{smathpar}
    (\E, \Theta, \langle s,i,op_\tau ::
    \sigma \rangle \pll \Sigma) \xrightarrow{\eff} (\E', \Theta',
    \langle s,i+1,\sigma \rangle \pll \Sigma)
  \end{smathpar}
  Cases:
  \begin{itemize}
    \item Case \rulelabel{EC}: Hypotheses:
      \begin{smathpar}
      \begin{array}{cr}
      \tau = {\sf EventuallyConsistent} & H\npp \\
      \auxred{\Theta}{(\E,\langle s,i,op \rangle)} {r} 
        {(\E',\eff)} & H\npp\\
      \end{array}
      \end{smathpar}
      Goal (\romannumeral 1) follows from $H1$ and
      Lemma~\ref{lem:auxred-wf}. Goal (\romannumeral 2) is the
      following: 
      \begin{smathpar}
      \begin{array}{cr}
        \E' \models \forall a,b. ~\hbo{a}{b} \wedge \vis{b}{\eff}
          \Rightarrow \vis{a}{\eff} & G\mpp\\
      \end{array}
      \end{smathpar}
      Let $\EffSoup' = \E'.\EffSoup$, $\Rvis' = \E'.\Rvis$, $\Rso' =
      \E'.\Rso$, and $\hboZ' =(\Rso' \cup \Rvis')^{+}$.  By inversion
      on $H1$, we get the following hypotheses:
      \begin{smathpar}
      \begin{array}{cr}
        \EffSoup' = \EffSoup \cup \{\eff\} & H\npp\\
        \visZ' = \Theta(r)\times\eff ~\cup~ \visZ & H\npp\\
        \Rso' = \EffSoup_{{\sf SessID}=s,\,{\sf SeqNo}<i}\times\eff ~\cup~ \Rso & H\npp\\
        \sameobjZ' = \EffSoup' \times \EffSoup'& H\npp\\
      \end{array}
      \end{smathpar}
       Since $\E'$ defines $\EffSoup'$ as the universe of values.
       Therefore, the goal can be rewritten:
      \begin{smathpar}
      \begin{array}{cr}
        \forall (a,b\in\EffSoup').
        \msentails{E'}{\De{\hbo{a}{b} \wedge
        \vis{b}{\eff}} \Rightarrow 
          \De{\vis{a}{\eff}}} & G\mpp\\
      \end{array}
      \end{smathpar}
      New hypotheses after \emph{intros}:
      \begin{smathpar}
      \begin{array}{cr}
        a \in \EffSoup' & H\npp\\
        b \in \EffSoup' & H\npp\\
      \end{array}
      \end{smathpar}
      And new goal:
      \begin{smathpar}
      \begin{array}{cr}
        \msentails{E'}{\De{\hbo{a}{b} \wedge
        \vis{b}{\eff}} \Rightarrow 
          \De{\vis{a}{\eff}}} & G\mpp\\
      \end{array}
      \end{smathpar}
      Since $(\mathcal{M} \models A \Rightarrow B) \Leftrightarrow
      (\mathcal{M} \models A \Rightarrow \mathcal{M} \models B)$, we
      prove $G0$ by proving: 
      \begin{smathpar}
      \begin{array}{cr}
        (\E' \models ~\hbo{a}{b} \wedge \vis{b}{\eff})
          \Rightarrow (\E' \models \vis{a}{\eff}) & G\mpp\\
      \end{array}
      \end{smathpar}
      After intros:
      \begin{smathpar}
      \begin{array}{cr}
        \E' \models ~\hbo{a}{b} \wedge \vis{b}{\eff} &
        H\npp\\
      \end{array}
      \end{smathpar}
      Since $\E'$ defines $\hboZ'$ and $\Rvis'$ as interpretations for
      $\hboZ$ and $\Rvis$ respectively, we have:
      \begin{smathpar}
      \begin{array}{cr}
          {{\sf hbo'}(a,b)} \wedge \Rvis'(b,\eff) & H\npp\\
      \end{array}
      \end{smathpar}
      And the goal is:
      \begin{smathpar}
      \begin{array}{cr}
        \Rvis'(a,\eff) & G\mpp\\
      \end{array}
      \end{smathpar}
      Inversion on $H9$:
      \begin{smathpar}
      \begin{array}{cr}
          {{\sf hbo'}(a,b)} & H\npp\\
          \Rvis'(b,\eff) & H\npp\\
      \end{array}
      \end{smathpar}
      Since $\eff$ is unique, from $H3$ and $H11$ we have the following:
      \begin{smathpar}
      \begin{array}{cr}
            b\in\Theta(r)& H\npp\\
      \end{array}
      \end{smathpar}
      Since local state (i.e., $\Theta(r)$) is causally consistent
      (lemma~\ref{lem:local-state-is-cc}) using $H10$ and $H12$ we
      derive the following:
      \begin{smathpar}
      \begin{array}{cr}
        a \in \Theta(r) & H\npp\\
      \end{array}
      \end{smathpar}
      Now, from $H3$ and $H13$, we deduce:
      \begin{smathpar}
      \begin{array}{cr}
        (a,\eff) \in \Rvis'& \\
      \end{array}
      \end{smathpar}
      which is what needs to be proven ($G4$).

    \item Case $\rulelabel{CC}$: Hypotheses:
      \begin{smathpar}
      \begin{array}{cr}
        \tau = {\sf CausallyConsistent} & H\npp \\
        \auxred{\Theta}{(\E,\langle s,i,op \rangle)} {r} 
          {(\E',\eff)} & H\npp\\
         \E'.\Rso(\eff) \subseteq \Theta(r)& H\npp\\
      \end{array}
      \end{smathpar}
     Goal (\romannumeral 1) follows from $H15$ and
     Lemma~\ref{lem:auxred-wf}. We now prove Goal (\romannumeral 2).
     Expanding the definition of $\ccc$, goal is the following:
      \begin{smathpar}
      \begin{array}{cr}
        \ccc = \forall a.~\hbo{a}{\eff} \Rightarrow
        \vis{a}{\eff} & G\mpp\\
      \end{array}
      \end{smathpar}
      Let $\EffSoup' = \E'.\EffSoup$, $\Rvis' = \E'.\Rvis$, $\Rso' =
      \E'.\Rso$, and $\hboZ' =(\Rso' \cup \Rvis')^{+}$. By inversion
      on $H15$, we get the following:
      \begin{smathpar}
      \begin{array}{cr}
        \EffSoup' = \EffSoup \cup \{\eff\} & H\npp\\
        \visZ' = \Theta(r)\times\eff ~\cup~ \visZ & H\npp\\
        \Rso' = \EffSoup_{{\sf SessID}=s,\,{\sf SeqNo}<i}\times\eff ~\cup~ \Rso & H\npp\\
        \sameobjZ' = \EffSoup' \times \EffSoup'& H\npp\\
      \end{array}
      \end{smathpar}
%     From $H16$ and $H19$, we have:
%     \begin{smathpar}
%     \begin{array}{cr}
%       \EffSoup_{{\sf SessID}=s,\,{\sf SeqNo}<i} \in \Theta(r) &
%       H\npp \\
%     \end{array}
%     \end{smathpar}
      By similar reasoning as with the previous case (\rulelabel{EC}),
      we get a context with following hypotheses:
      \begin{smathpar}
      \begin{array}{cr}
        a \in \EffSoup' & H\npp\\
        {\hboZ'(a,\eff)}  & H\npp\\
      \end{array}
      \end{smathpar}
      And the goal is the following:
      \begin{smathpar}
      \begin{array}{cr}
        \Rvis'(a,\eff) & G\mpp\\
      \end{array}
      \end{smathpar}
      Since \emph{happens-before} is transitive, by inversion on
      $H23$, we get two cases:
      \begin{itemize}
        \item SCase 1: Hypotheses:
        \begin{smathpar}
        \begin{array}{cr}
          (\Rso' \cup \Rvis')(a,\eff) & H\npp\\
        \end{array}
        \end{smathpar}
        Inversion on $H24$ leads to two subcases. In one case, we
        assume $\Rvis'(a,\eff)$ and try to prove the goal $G6$. The
        proof for this case mimics the proof for Case \rulelabel{EC}.
        Alternatively, in second case, we assume:
        \begin{smathpar}
        \begin{array}{cr}
          \Rso'(a,\eff) & H\npp\\
        \end{array}
        \end{smathpar}
        and prove $G6$. From $H25$ and $H16$, we infer:
        \begin{smathpar}
        \begin{array}{cr}
          a \in \Theta(r) & H\npp\\
        \end{array}
        \end{smathpar}
        Now, from $H26$ and $H18$ we know:
        \begin{smathpar}
        \begin{array}{cr}
          (a,\eff) \in \Rvis' & \\
        \end{array}
        \end{smathpar}
        which is the goal ($G6$).

        \item SCase 2: Hypotheses (after abbreviating the occurance
        of$(\Rso' \cup \Rvis')^{+}$ as {\sf hbo'}):
        \begin{smathpar}
        \begin{array}{cr}
          \exists(c\in\EffSoup').{\sf hbo'}(a,c) \wedge (\Rso' \cup
          \Rvis')(c,\eff) & H\npp\\
        \end{array}
        \end{smathpar}
        Inverting $H27$, followed by expanding $\Rso' \cup \Rvis'$:
        \begin{smathpar}
        \begin{array}{cr}
          c\in\EffSoup' & H\npp\\
           {\sf hbo'}(a,c) \wedge (\Rso'(c,\eff) \vee \Rvis'(c,\eff))& H\npp\\
        \end{array}
        \end{smathpar}
        Inverting the disjunction in $H29$, we get two cases:
        \begin{itemize}
          \item SSCase R: Hypothesis is
          \begin{smathpar}
          \begin{array}{cr}
             {\sf hbo'}(a,c) \wedge \Rvis'(c,\eff)& H\npp\\
          \end{array}
          \end{smathpar}
          Observe that hypothesis $H30$ and current goal ($G6$) are
          same as hypothesis $H9$ and goal ($G4$) in Case
          \rulelabel{EC}. The proof for this SSCase is also the same.

          \item SSCase L: Hypothesis is
          \begin{smathpar}
          \begin{array}{cr}
             {\sf hbo'}(a,c) \wedge \Rso'(c,\eff)& H\npp\\
          \end{array}
          \end{smathpar}
          Inverting $H31$:
          \begin{smathpar}
          \begin{array}{cr}
             {\sf hbo'}(a,c)& H\npp\\
             \Rso'(c,\eff)& H\npp\\
          \end{array}
          \end{smathpar}
          From $H33$ and $H16$, we infer:
          \begin{smathpar}
          \begin{array}{cr}
            c \in \Theta(r) & H\npp \\
          \end{array}
          \end{smathpar}
          Since local state is causally consistent
          (lemma~\ref{lem:local-state-is-cc}), from $H32$ and $H34$,
          we have:
          \begin{smathpar}
          \begin{array}{cr}
            a \in \Theta(r) & H\npp\\
          \end{array}
          \end{smathpar}
           Proof follows from $H35$ and $H18$.
        % End of SSCase
        \end{itemize}
      % End of SCases
      \end{itemize}

    \item Case \rulelabel{SC}: Hypotheses:
      \begin{smathpar}
      \begin{array}{cr}
        \tau = {\sf StronglyConsistent} & H\npp \\
        \auxred{\Theta} {(\E,\langle s,i,op \rangle)} {r}
        {(\E',\eff)} & H\npp\\ 
        \E.A \subseteq \Theta(r) & H\npp\\
        \dom(\Theta') = \dom(\Theta) & H\npp\\
        \forall r'\in \dom(\Theta'). \Theta'(r') = \Theta(r') \cup
          \{\eff\} & H\npp\\
      \end{array}
      \end{smathpar}
      Let $\EffSoup' = \E'.\EffSoup$, $\Rvis' = \E'.\Rvis$, $\Rso' =
      \E'.\Rso$, and $\hboZ' =(\Rso' \cup \Rvis')^{+}$. Inversion on
      $H37$ gives: 
      \begin{smathpar}
      \begin{array}{cr}
        \EffSoup' = \EffSoup \cup \{\eff\} & H\npp\\
        \visZ' = \Theta(r)\times\eff ~\cup~ \visZ & H\npp\\
        \Rso' = \EffSoup_{{\sf SessID}=s,\,{\sf SeqNo}<i}\times\eff ~\cup~ \Rso & H\npp\\
        \sameobjZ' = \EffSoup' \times \EffSoup'& H\npp\\
      \end{array}
      \end{smathpar}
       Goal (\romannumeral 1) follows from $H36$ and
       Lemma~\ref{lem:auxred-wf}. Expanding the definition of $\scc$,
       goal is the following:
      \begin{smathpar}
      \begin{array}{cr}
        \forall a.~\sameobj{a}{\eff} \Rightarrow \vis{a}{\eff} 
          ~\vee~ \vis{\eff}{a} ~\vee~ a = \eff & G\mpp\\
      \end{array}
      \end{smathpar}
      By similar reasoning as with the \rulelabel{EC} case,
      we get a context with following hypotheses:
      \begin{smathpar}
      \begin{array}{cr}
        a \in \EffSoup' & H\npp\\
        {\sameobjZ'(a,\eff)}  & H\npp\\
      \end{array}
      \end{smathpar}
      And the goal is the following:
      \begin{smathpar}
      \begin{array}{cr}
        \vis{a}{\eff} ~\vee~ \vis{\eff}{a} ~\vee~ a = \eff & G\mpp\\
      \end{array}
      \end{smathpar}
      From $H40$ and $H44$, we know that either $a \in \EffSoup$ or $a
      = \eff$. If $a = \eff$, then $G8$ follows trivially. On the
      other hand, when $a \in \EffSoup$, from $H37$ we know that $a
      \in \Theta(r)$. From $H41$, we derive:
      \begin{smathpar}
      \begin{array}{cr}
        a \in \visZ'(\eff)& \\
      \end{array}
      \end{smathpar}
      Which means first disjunct of $G8$ is valid, proving the
      theorem.
  % End of all cases    
  \end{itemize}

\end{proof}

\begin{corollary}[Soundness]
\label{thm:soundness}
For every well-formed execution state $\E$, and contract class $\tau
\in \{{\sf ec},{\sf cc},{\sf sc}\}$, and for every consistency
contract $\cv$ in the contract class $\tau$, If:
\begin{smathpar}
(\E, \Theta, \langle s,i,op_\tau :: 
\sigma \rangle \pll \Sigma) \xrightarrow{\eff} (\E', \Theta',
\langle s,i+1,\sigma \rangle \pll \Sigma)
\end{smathpar}
then (\romannumeral 1) $\E'$ is well-formed, and (\romannumeral 2)
$\E' \models \cv[\eff/\cureff]$
\end{corollary}
\begin{proof}
Follows from Theorems~\ref{thm:classification-sound} and
\ref{thm:weak-soundness}
\qed
\end{proof}
